---
title: "Dissemination Areas to target"
output: html_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE)

```

# Introduction

This piece looks at Dissemination Areas to target. 

## Remaining items

1. Incumbent results  
2. Regression modelling to identify features strongly correlated with candidate success  
3. Characterize each subdivision

# Hyper parameters

Set hyperparameters below. 

```{r}

# set councillor
councillor <- 'Olawoye Lekan'

# set year
year <- 2014

# set percentage_from_top_candidate
percentage_from_top_candidate <- 0.05

```

# Administrative

Load packages

```{r}

# data manipulation packages
library(dplyr)
library(tidyr)

# visualization packages
library(ggplot2)
library(plotly)

# mapping packages
library(rgdal)
library(sp)
library(scales)


# download hyperparameters
source('C:\\Users\\fmqai\\Documents\\Election_Results\\Hyper_parameters.R')

```

# Themes

```{r}

# set colors

color1 = '#C21032'
color2 = '#F8E45F'
color3 = '#C8CE86'
color4 = '#424441'
color5 = '#728D6E'

text_color1 = 'white'
text_color2 = 'grey30'

# Font

# extrafont::font_import(pattern="Arial", prompt = FALSE) # load all fonts

font1 = 'Tw Cen MT'
font_size1 = '11'
font_style1 = 'bold'

# set blank theme

theme_blank <- theme(
  axis.line = element_blank()
  ,axis.text.x = element_blank()
  ,axis.text.y = element_blank()
  ,axis.ticks = element_blank()
  ,axis.title.x = element_blank()
  ,axis.title.y = element_blank()
  ,legend.position = "none"
  ,panel.background = element_blank()
  ,panel.border = element_blank()
  ,panel.grid.major = element_blank()
  ,panel.grid.minor = element_blank()
  ,plot.background = element_blank()
  )

```

# Load data

```{r}

# read in election results data
if (!exists('election_results_raw')) {
  temp_path <- 'C:\\Users\\fmqai\\OneDrive\\Projects\\Wards\\R generated data\\Election_results.csv'
  election_results_raw <- read.csv(file = temp_path, header = TRUE, stringsAsFactors = FALSE)}

# read in ward/subdivision data
if (!exists('demographics_sd_raw')) {
  temp_path <- 'C:\\Users\\fmqai\\OneDrive\\Projects\\Wards\\R generated data\\Census_data\\Subdivision.csv'
  demographics_sd_raw <- read.csv(file = temp_path, header = TRUE, stringsAsFactors = FALSE)}

# read in dissemination area data
# if (!exists('demographics_da_raw')) {
#   temp_path <- 'C:\\Users\\fmqai\\OneDrive\\Projects\\Wards\\R generated data\\Census_data\\Dissemination_area.csv'
#   demographics_da_raw <- read.csv(file = temp_path, header = TRUE, stringsAsFactors = FALSE)}

# read in master_linkage
if (!exists('geo_linkage_raw')) {
  temp_path <- 'C:\\Users\\fmqai\\OneDrive\\Projects\\Wards\\R generated data\\Shapefiles\\Linking_Shapefiles\\Master_linkage.csv'
  geo_linkage_raw <- read.csv(file = temp_path, header = TRUE, stringsAsFactors = FALSE)}

# read in language classification file Farooq produced
if (!exists('language_classification_raw')) {
  temp_path <- 'C:\\Users\\fmqai\\OneDrive\\Projects\\Wards\\R Inputs\\Census Data\\Meta\\Language_classification.csv'
  language_classification_raw <- read.csv(file = temp_path, header = TRUE, stringsAsFactors = FALSE)}

# read in subdivision aggregating classification file Farooq produced
if (!exists('aggregating_sd_raw')) {
  temp_path <- 'C:\\Users\\fmqai\\OneDrive\\Projects\\Wards\\R Inputs\\Election Results\\Meta\\Aggregating_subdivisions.csv'
  aggregating_sd_raw <- read.csv(file = temp_path, header = TRUE, stringsAsFactors = FALSE)}

# read in ward/subdivision shapefile
if (!exists('ward_shp')){
  temp_path <- paste0(input_location, "Shapefiles\\Voting Subdivisions")
  ward_shp <- readOGR(dsn = temp_path, layer = "VOTING_SUBDIVISION_2014_WGS84")}

# read in dissemination area shapefile
if (!exists('da_shp')){
  temp_path <- paste0(input_location, "Shapefiles\\Dissemination Areas")
  da_shp <- readOGR(dsn = temp_path, layer = "gda_000b11a_e")}


```

```{r}

# Based on the hyperparamaters set above, define other attributes

wards <- unique(election_results_raw[election_results_raw$Candidate==councillor
                                 & election_results_raw$Year==year, ]$Ward)

subdivisions <- unique(election_results_raw[election_results_raw$Candidate==councillor
                                        & election_results_raw$Year==year, ]$Subdivision)


```

# `r Year` `r councillor` Results

## Ward results

```{r}

# We will use this dataframe for all election results
# This way we won't have to aggregate repeatedly
election_results <- election_results_raw %>%
  left_join(select(aggregating_sd_raw, AreaName, Associated_Subdivision, Associated_AreaName), 
            'AreaName') %>%
  mutate(AreaName = ifelse(!is.na(Associated_AreaName), 
                           Associated_AreaName, 
                           AreaName),
         Subdivision = ifelse(!is.na(Associated_Subdivision), 
                              Associated_Subdivision, 
                              Subdivision)
         ) %>%
  # summarise 
  group_by(AreaName, Year, Ward, Candidate, Subdivision) %>%
    summarise(Votes = sum(Votes, na.rm = TRUE)) %>%
  ungroup() %>%
  # set correct format for column
  mutate(AreaName = as.character(AreaName)) %>%
  # select final column in following order
  select(AreaName, Year, Ward, Candidate, Subdivision, Votes)

```


```{r}

# create results for selected councillor, year 
temp <- election_results %>%
  filter(Year %in% year) %>%
  filter(Ward %in% wards) %>%
  # calculate total votes for each Candidate
  group_by(Year, Ward, Candidate) %>%
    summarise(Votes = sum(Votes, na.rm = TRUE)) %>%
  ungroup() %>%
  # calculate percentage of votes
  mutate(Percentage = Votes/sum(Votes, na.rm = TRUE)) %>%
  # calculate percentage from top candidate
  mutate(Percentage_max = max(Percentage, na.rm = TRUE), 
         Percentage_difference_from_max = Percentage_max - Percentage
         ) %>%
  # create labels
  mutate(# Percentage label, 
         Percentage_label = paste0(round(Percentage*100, 0), '%'), 
         # calculate location of Percentage_label
         Percentage_loc = Percentage/2, 
         # votes label 
         Votes_label = paste0(format(Votes, big.mark = ',')), 
         # calculate location of Votes_label
         Votes_loc = Percentage + 0.01,
         # Percentage from top label
         Percentage_difference_from_max_label = paste0(round(-Percentage_difference_from_max*100, 0), '%'))

# order candidates by votes
temp$Candidate <- factor(x = temp$Candidate, 
                         levels = temp$Candidate[order(temp$Votes, decreasing = TRUE)])
  
  
# create plot
temp_width <- 0.7

p <- ggplot(data = temp, mapping = aes(x = Candidate, y = Percentage)) + 
  geom_bar(stat = 'identity', fill = color1, width = temp_width) + 
  # percentage labels in center of bars
  geom_text(mapping = aes(label = Percentage_label, 
                          y = Percentage_loc), 
            color = text_color1, 
            fontface = font_style1) +
  # votes label on top of bars 
  geom_text(aes(label = Votes_label, y = Votes_loc), color = text_color2, family = 'Helvetica', fontface = font_style1) + 
  # difference from our candidate and top candidate 
  geom_segment(aes(x = as.numeric(Candidate) + temp_width/3, y = Percentage_max,
                   xend = as.numeric(Candidate) + temp_width/3, yend = Percentage), 
               color = ifelse(temp$Candidate==councillor, color2, NA), 
               size = 1,
               lineend = "square", 
               arrow = arrow(type = 'closed') 
               ) +
  # difference label 
  geom_text(aes(label = ifelse(temp$Candidate==councillor, 
                               Percentage_difference_from_max_label, 
                               NA),
                x = Candidate,
                y = Percentage_difference_from_max/2 + Percentage, 
                hjust = -temp_width*2.5,
                fontface = font_style1
                ), 
            color = color2) +
  theme_blank +
  theme(axis.text.x = element_text(family = font1, 
                                   face = font_style1,
                                   size = font_size1,
                                   color = text_color2))

# set winning candidate
winning_candidate <- as.character(temp[which.max(temp$Votes), ]$Candidate)

# print plot
p

  
```

## Subdivision results

```{r}

# create temporary dataset linking AreaNames to AreaIDs
temp_geo_linkage <- select(geo_linkage_raw, AREA_NAME, AREA_ID) %>%
  mutate(AREA_NAME = as.character(AREA_NAME), 
         AREA_ID = as.character(AREA_ID)) %>%
  distinct() %>%
  filter(!is.na(AREA_NAME))

# extract centroids
sd_centroids <- data.frame(rgeos::gCentroid(ward_shp, byid=TRUE)
                             , ward_shp@data
                             , stringsAsFactors = FALSE) %>%
  rename(Label_x = x
         , Label_y = y) %>%
  mutate(AREA_ID = as.character(AREA_ID)) %>%
  select(AREA_ID, Label_x, Label_y)

# create results datasets
councillor_election_results <- election_results %>%
  filter(Year %in% year) %>%
  filter(Ward %in% wards) %>%
  # summarize data 
  group_by(Year, Ward, Candidate, Subdivision, AreaName) %>%
    summarise(Votes = sum(Votes, na.rm = TRUE)) %>%
  ungroup() %>%
  # calculate percentage of votes
  group_by(Year, Ward, Subdivision) %>%
    mutate(Percentage = Votes/sum(Votes, na.rm = TRUE)) %>%
  ungroup() %>%
  # calculate percentage from top candidate
  group_by(Year, Ward, Subdivision) %>%
    mutate(Percentage_from_top = max(Percentage, na.rm = TRUE) - Percentage) %>%
  ungroup() %>%
  # Candidate rank in subdivision
  group_by(Year, Ward, Subdivision) %>%
    mutate(Rank = as.factor(dense_rank(desc(Percentage)))) %>%
  ungroup() %>%
  # add AREA_ID
  left_join(temp_geo_linkage, 
            c('AreaName'='AREA_NAME')) %>%
  # create labels column
  mutate(Label = paste0(round(-Percentage_from_top*100, 0), '%')) %>%
  # add Label locations
  left_join(sd_centroids, 'AREA_ID') %>%
  # select final columns
  select(Year, AREA_ID, Ward, Subdivision, Candidate, Votes, Percentage,
         Percentage_from_top, Rank, Label, Label_x, Label_y)

```


```{r}

# get AREA_ID for ward in question
ward_AREA_IDs <- unique(councillor_election_results[councillor_election_results$Ward %in% wards, ]$AREA_ID)

# filter for councillor in question
temp_councillor_election_results <- councillor_election_results %>%
  filter(Candidate==councillor)

# convert to shapefile to dataframe for ggplot2
temp_shp <- fortify(ward_shp, region = 'AREA_ID') %>%
  # filter for the ward we want
  filter(id %in% ward_AREA_IDs) %>%
  # add data
  left_join(temp_councillor_election_results, c("id"='AREA_ID'))


# create plot
p <- ggplot(temp_shp, aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = Rank)
                 , color = 'black'
                 , alpha = 1
               ) +
  geom_text(aes(label = Subdivision, 
                x = Label_x, 
                y = Label_y), 
            color = ifelse(temp_shp$Rank==4, text_color2, text_color1)
            ) +
  scale_fill_brewer(palette = 'Blues',
                    direction = -1, 
                    name = 'Subdivision\nRank',
                    guide = guide_legend(reverse = FALSE)) +
  coord_fixed(1.3) +
  theme_blank + 
  theme(#panel.background = element_rect(fill = 'black', colour = 'black'), 
        plot.background = element_rect(fill = 'black', colour = 'black'),
        legend.background = element_rect(fill = 'black', color = 'black'),
        legend.text = element_text(color = 'white'),
        legend.title = element_text(color = 'white'),
        legend.position = 'right')

# print plot
# ggplotly(p)
p

```

```{r}

temp_councillor_election_results %>%
  mutate(Percentage = paste0(round(Percentage*100, 0), '%')) %>%
  select(Year, Ward, Subdivision, Candidate, Votes, Percentage, Rank)

```

## Swing Subdivisions

```{r}

# get AREA_ID for ward in question
ward_AREA_IDs <- unique(councillor_election_results[councillor_election_results$Ward %in% wards, ]$AREA_ID)

# filter for councillor in question
temp_councillor_election_results <- councillor_election_results %>%
  filter(Candidate==councillor)

# convert to shapefile to dataframe for ggplot2
temp_shp <- fortify(ward_shp, region = 'AREA_ID') %>%
  # filter for the ward we want
  filter(id %in% ward_AREA_IDs) %>%
  # add data
  left_join(temp_councillor_election_results, c("id"='AREA_ID'))

# create plot
p <- ggplot(temp_shp, aes(x = long, y = lat, group = group)) + 
  geom_polygon(aes(fill = Percentage_from_top)
                 , color = 'black'
                 , alpha = 1
               ) +
  geom_text(aes(label = Subdivision, 
                x = Label_x, 
                y = Label_y), 
            color = ifelse(temp_shp$Percentage_from_top<0.1, text_color1, text_color2)
            ) +
  scale_fill_distiller(palette = 'Blues', 
                       name = 'Percentage from Subdivision\nWinning Candidate',
                       guide = guide_legend(reverse = FALSE)) +
  coord_fixed(1.3) +
  theme_blank + 
  theme(#panel.background = element_rect(fill = 'black', colour = 'black'), 
        plot.background = element_rect(fill = 'black', colour = 'black'),
        legend.background = element_rect(fill = 'black', color = 'black'),
        legend.text = element_text(color = 'white'),
        legend.title = element_text(color = 'white'),
        legend.position = 'right')

# print plot
# ggplotly(p)
p

```

```{r}

temp_councillor_election_results %>%
  mutate(Percentage_from_top = paste0(round(Percentage_from_top*100, 0), '%')) %>%
  select(Year, Ward, Subdivision, Candidate, Percentage_from_top)

```

# Characterizing subdivisions

```{r}

# We will use this dataframe for all demographics at subdivision level
# This way we won't have to aggregate repeatedly

# create temporary dataset linking AreaNames to AreaIDs
temp_geo_linkage <- geo_linkage_raw %>%
  select(AREA_NAME, AREA_ID) %>%
  mutate(AREA_NAME = as.character(AREA_NAME), 
         AREA_ID = as.character(AREA_ID)) %>%
  rename(AreaName = AREA_NAME, 
         AreaID = AREA_ID) %>%
  distinct() %>%
  filter(!is.na(AreaName))

# aggregation dataframe idetifying which subdivision 'voting places' are associated with
aggregating_sd <- aggregating_sd_raw %>%
  mutate(AreaName = as.character(AreaName)) %>%
  left_join(temp_geo_linkage, 'AreaName') %>%
  select(AreaID, AreaName, Associated_Subdivision, Associated_AreaName)

demographics_sd <- demographics_sd_raw %>%
  rename(AreaID = AREA_ID) %>%
  mutate(AreaID = as.character(AreaID)) %>%
  left_join(aggregating_sd, 'AreaID') %>%
  mutate(AreaName = ifelse(!is.na(Associated_AreaName), 
                           Associated_AreaName, 
                           AreaName),
         Subdivision = ifelse(!is.na(Associated_Subdivision), 
                              Associated_Subdivision, 
                              Subdivision)
         ) %>%
  # summarise 
  group_by(AreaName, Year, Ward, Candidate, Subdivision) %>%
    summarise(Votes = sum(Votes, na.rm = TRUE)) %>%
  ungroup() %>%
  # set correct format for column
  mutate(AreaName = as.character(AreaName)) %>%
  # select final column in following order
  select(AreaName, Year, Ward, Candidate, Subdivision, Votes)


```


```{r}

# Get demographics data into order

# Age

temp_youth_categories <- c("0 to 4 years", 
                           "5 to 9 years",
                           "10 to 14 years",
                           "15 to 19 years",
                           "20 to 24 years", 
                           "25 to 29 years")

temp_age <- demographics_sd %>%
  filter(Topic=='Age characteristics') %>%
    # before aggregating subdvisions, we need to ensure "voting places"" identify with a subdivision in our shapefile
  left_join(aggregating_sd, c('Ward', 'Subdivision')) %>%
  mutate(Subdivision = ifelse(!is.na(Associated_Subdivision), Associated_Subdivision, Subdivision)) %>%
  mutate(Category = ifelse(Characteristic %in% temp_youth_categories, 
                           'Youth', 
                           'Non_Youth')) %>%
  group_by(AREA_ID, Category) %>%
    summarise(Total = sum(Total, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(AREA_ID) %>%
    mutate(Percentage = Total/sum(Total, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-Total) %>%
  spread(Category, Percentage)

# Language

# n.i.e. = not included elsewhere 

temp_language <- demographics_sd_raw %>%
  filter(Topic=='Detailed language spoken most often at home') %>%
    # before aggregating subdvisions, we need to ensure "voting places"" identify with a subdivision in our shapefile
  left_join(aggregating_sd, c('Ward', 'Subdivision')) %>%
  mutate(Subdivision = ifelse(!is.na(Associated_Subdivision), Associated_Subdivision, Subdivision)) %>%
  left_join(select(language_classification, Characteristic, Classification), 'Characteristic') %>%
  mutate(Classification = ifelse(Classification=='Other', 'Other language', Classification),  
         Category = Classification) %>%
  group_by(AREA_ID, Category) %>%
    summarise(Total = sum(Total, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(AREA_ID) %>%
    mutate(Percentage = Total/sum(Total, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-Total) %>%
  spread(Category, Percentage)

# Dwellings

temp_dwelling <- demographics_sd_raw %>%
  filter(Topic=='Household and dwelling characteristics') %>%  # before aggregating subdvisions, we need to ensure "voting places"" identify with a subdivision in our shapefile
  left_join(aggregating_sd, c('Ward', 'Subdivision')) %>%
  mutate(Subdivision = ifelse(!is.na(Associated_Subdivision), Associated_Subdivision, Subdivision)) %>%
  within({
    Category = Characteristic
    Category[grepl(x = Characteristic, '.*Apartment.*')] = 'Apartment'
    Category[Characteristic %in% c('Movable dwelling', 'Other dwelling', 'Other single-attached house')] = 'Other dwelling'
  }) %>%
  group_by(AREA_ID, Category) %>%
    summarise(Total = sum(Total, na.rm = TRUE)) %>%
  ungroup() %>%
  group_by(AREA_ID) %>%
    mutate(Percentage = Total/sum(Total, na.rm = TRUE)) %>%
  ungroup() %>%
  select(-Total) %>%
  spread(Category, Percentage)

# One table to rule them all! LOTR

temp_demographics <- temp_age %>%
  left_join(temp_language, 'AREA_ID') %>%
  left_join(temp_dwelling, 'AREA_ID') %>%
  mutate(AREA_ID = as.character(AREA_ID))

colnames(temp_demographics) <- gsub(x = colnames(temp_demographics), pattern = ' |-', replacement = '_')


```


```{r}

all_data <- councillor_election_results %>%
  left_join(temp_demographics, 'AREA_ID') 


  

```
